How the Practice of TDD Influences \\Class Design in Object-Oriented Systems: \\Patterns of Unit Tests Feedback

### 2013-06-03 jss-2013/paper.tex
primeira revisao das modificacoes sugeridas pelo journal

\begin{abstract}

	Despite the fact that Test-Driven Development (TDD) appears to be a software testing practice, 
	many developers affirm that the practice influences on class design. This study aimed
	to better understand the effects of TDD and how the practice influences developers
	during class design on object-oriented systems. 
	We conducted an essentially qualitative exploratory study in which participants
	were invited to implement some exercises using TDD and, based on the data gathered,
	we raised details of how the practice influenced design decisions from the
	participants using interviews.
	We noticed that the practice of TDD drives
	developers during class design by means of constant feedback about its quality. This study
	also named and catalogued feedback patterns perceived by the developers.

@@

To understand these reasons, it is necessary to conduct a research on the real world, which
implies on a balance between control and realism. A realistic situation is usually
complex and non deterministic, which makes the comprehension of the phenomenon much
harder. On the other hand, increasing the control of the experiment reduces the
degree of realism, which can take the real factors of influence out of the study
scope \cite{guidelines-case-study}.

Based on the fact that the software development process involves many
human factors and is totally sensitive to the context that it belongs to,
we conducted an essentially qualitative exploratory study in which professionals
from the industry were invited to implement some prepared exercises using TDD and,
based on the gathered data, we investigated the details of how the practice influenced
the participants' class design decisions by means of interviews.

@@

It is a known fact that class design tends to lose quality during the software evolution \cite{evolution-lehman}.
However, it is hard to discuss quality in terms of class design. In this study context,
we used the SOLID principles, detailed in Martin's work \cite{bob-martin}.

@@

@@

Many empirical studies have been conducted to evaluate the effects of TDD.
In most of them, the effects of the practice in the class design are not taken
into account, and only the effect on external quality is measured.
Contrary to what this study proposes, many of them chose
to have a bigger control in the experiment. They were also conduted of the academic
environment, with students that belong to different computer courses.

@@

However, studies that analyze TDD from the class design point of view does not
come to conclusive results; many of them even affirm that the effects of TDD
are not so different from those teams that do not practice it. Even the Janzen's PhD
thesis was inconclusive regarding the influence of the practice in coupling
and cohesion \cite{janzen-phd}.

@@

Participants from different software development companies in the Brazilian
industry were selected. The participants' profiles are discussed in sub-section
\ref{sec:planejamento-participantes}. All of them were invited to implement
a few problems using Java in a defined timeframe. Participants practiced
TDD in one problem and did not practice in the other. The problems,
as well as the instruction of using or not TDD, were
randomized in order to reduce bias.

@@

They evaluated the code based on three different categories: \textit{Simplicity}, \textit{Testability},
and \textit{Class Design Quality}. In each category, specialists scored from 1 (bad) to
5 (good) or choose not to evaluate that exercise. As a few participants did not finish the
implementation, specialists were warned to evaluate the intention of the participant and not
only the produced code.
To reduce the bias, the specialists did not know which code was written with TDD.

@@

We had 25 participants from 6 different companies. The participants, in their majority,
were not experienced in TDD. 40\% of them affirmed to use TDD for no more than a year.
52\% of them practiced TDD from 1 to 3 years. Only 4\% practiced TDD from 3 to 4
years, and no participant was more experienced than that.

The numbers were a little different when talking about their experience in software
development. 24\% of them worked with software development for the last 4 or 5 years.
28\% of them worked from 6 to 10 years. Only 20\% of them developed software for less
than 2 years. 64\% of the participants use Java. However, 36\% of them do not use
Java in their daily work. All of them know JUnit and 64\% use mock objects 
\footnote{Mock objects are objects created during a unit test. They mock the behavior 
of a concrete object. Usually they are used to isolate the unit test from other classes. 
More information about it can be found on \cite{mocks}.}
during their development activities.
Only 12\% have never heard about mock objects. When talking about object-orientation,
in the open question, most of the participants affirmed to have a good experience
on it. Some of them even affirmed to master the subject. Only a few of them affirmed
to have basic knowledge about OO.

@@

The elevated number of participants that do not use Java in their daily work is consequence
of a company that uses PHP. However, we checked that, although they do not use Java constantly,
they did not have troubles during the implementation of the exercises.

@@

We proposed four problems that should be implemented by the participants
\footnote{The exercises can be found at \url{http://gist.github.com/3024328}. Last access on July, the 30th, 2012.}. 
The goal was to simulate recurrent class design problems in several software projects.
In Table \ref{tab:problemas-exercicios}, we present the relation of a bad implementation of
the exercises and class design principles that were not followed.

@@

\begin{table}
	\centering
	\begin{tabular}{| p{2cm} | p{4cm} | p{2.5cm} | }
		\hline
		\textbf{Exercise} & \textbf{Design Smell} & \textbf{Principles to}\\
		& & \textbf{be Followed}\\

@@

		Exercise 1 & Rigidity, Needless Complexity & SRP, OCP \\
		Exercise 2 & Fragility, Viscosity, Immobility & SRP, DIP, OCP \\
		Exercise 3 & Rigidity, Fragility & SRP\\
		Exercise 4 & Fragility, Viscosity, Immobility & OCP, SRP, DIP \\

@@

@@

The statistical test used was the Wilcoxon. It is a non parametric hypothesis test
used to compare two samples and to check whether there is a difference in the meaning
of both samples. Therefore, we used Wilcoxon to compare the meaning between
the code metrics in code produced with and without TDD. The significance level
used was the standard (0.05).

@@

In Table \ref{metricas-industria}, we show the \textit{p-values} found for
the differences between codes produced with and without TDD. By the numbers,
we observe that there was no significant differences in cyclomatic complexity
and efferent coupling.
Lack of cohesion of methods was different in two exercises (1 and 4). A difference
also appeared on the quantity of lines in a method (exercise 4) and quantity of
methods (exercise 1).

@@

\begin{table*}
	\centering
	\begin{tabular}{ | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
		\hline
		\textbf{Exercise} & \textbf{Cyclomatic Complexity} & \textbf{Efferent coupling} & \textbf{Lack of Cohesion of Methods} & \textbf{Number of lines per method} 
		& \textbf{Quantity of methods per class} \\
		\hline
		Exercise 1 &	0.8967	&	0.6741 &	\cellcolor[gray]{0.8}2.04E-07* &	0.4962 &	\cellcolor[gray]{0.8}2.99E-06* \\
		Exercise 2	& 0.7868	&	0.7640 &	0.06132 &	0.9925 &	0.7501 \\
		Exercise 3	& 0.5463	&	0.9872 &	0.5471 &	0.7216 &	0.3972\\
		Exercise 4	& 0.2198	&	0.1361 &	\cellcolor[gray]{0.8}0.04891* &	\cellcolor[gray]{0.8}0.0032* &	0.9358\\
		\hline
		All together &	0.8123	&	0.5604 &	0.3278 &	0.06814 &	0.5849\\
		\hline
	\end{tabular}
	\caption{\textit{P-values} found for the difference between code produced with and without TDD}
	\label{metricas-industria}
\end{table*}

@@

%% ------------------------------------------------------------------------- %%
\subsection{Specialists}

@@

The values presented above corroborate with many of the related work.
Apparently TDD does not influence the code to the point of changing
the metric values for coupling, cohesion, and simplicity. However, this is
incoherent to the industry's common feeling of the positive effects of TDD on class design.
As stated, in this study we conducted a qualitative analysis to understand
how this influence goes from the point of view of the developers.

@@

Different from what was expected, the majority of the participants affirmed that the
practice of TDD would not change the class design produced by them during the exercises
in which they did not use the practice.
The main justification was that experience and previous knowledge on object-orientation
guided them during the class design. All participants agreed that a developer with no
knowledge in any of the mentioned areas would create a good class design only by practicing
TDD.

@@

The majority of the participants also commented that one difference that is
perceived when they practice TDD is the constant feedback. In traditional testing,
the time between the production code writing and test code writing is too long.
When practicing TDD, developers are forced to write the test first, and receive all
feedback a test can provide sooner.

@@

Also, the existence of the word \textit{"if"} in the name of the test is another
pattern. Tests that contain a name like that usually indicate the existence
of an "if" in the production code. Those many conditions might be refactored and,
through the use of polymorphism, eliminated. We call it \textbf{Conditional in Test's name}.

@@

		\textbf{Pattern} & \textbf{Class Design Smell} & \textbf{Not Followed Principles}\\

@@

@@

The proposed exercises were small compared to a real project. However, all exercises
contain localized class design problems. Once this study evaluates the effects
of TDD on class design, we believe the exercises can simulate real design problems
satisfactorily.
Besides, at the end of the implementation, participants answered a question about the
similarity between the exercises and real world problems. All of them affirmed that
the problems found in the exercises were very similar to the ones they deal with in real world.

@@

Some participants did not finish their implementation. This can influence the
quantitative analysis, as a design that would be complex at the end, looks simple
to an automated code metric.

@@

%% ------------------------------------------------------------------------- %%
\section{Conclusion}
\label{sec:conclusion}
%\section{Conclusão}

@@

In below subsections, we answer each of the questions raised by this study.

@@

The practice of TDD \textbf{can} affect the creating process of class design. However,
unlike the comments made by the industry, \textbf{the TDD practice cannot guide the
developer to a good class design automatically}; the experience and knowledge of
the developer are essential to develop oriented object software.

@@

\textbf{Therefore, this is the way how the practice guides the developer to
a better class design; giving constant feedback about potential issues in
class design. It is a developer's task to realize these issues and, 
improve the design. }

@@

%% ------------------------------------------------------------------------- %%
\begin{thebibliography}{9}

@@

	\bibitem{TDDByExample}
	------, \emph{Test-Driven Development By Example}, 1st~ed.\hskip 1em plus 0.5em
	  minus 0.4em\relax Addison-Wesley Professional, 2002.

	\bibitem{wambler-survey-agile}
	S.~Wambler, ``How agile are you? 2010 survey results,''
	  \url{http://www.ambysoft.com/surveys/howAgileAreYou2010.html}, 2010, Último
	  acesso em 28/10/2010.

	\bibitem{versionone-2012}
	V.~One, ``State of agile development survey results,''
	  \url{http://www.versionone.com/state_of_agile_development_survey/11/}, 2012,
	  Último acesso em 29/02/2012.

@@

	\bibitem{GOOS}
	S.~F. e~Nat~Pryce, \emph{Growing Object-Oriented Software, Guided by Tests},
	  1st~ed.\hskip 1em plus 0.5em minus 0.4em\relax Addison-Wesley Professional,
	  2009.

@@

	\bibitem{aniche-wbma}
	M.~Aniche, T.~Ferreira, and M.~Gerosa, ``What concerns beginner test-driven
	  development practitioners: A qualitative analysis of opinions in an agile
	  conference,'' 2011.

@@

	\bibitem{langr}
	J.~Langr, ``Evolution of test and code via test-first design,''
	  \url{http://www.objectmentor.com}, 2001, Último acesso em 01/03/2011.

@@

	\bibitem{josefsson}
	M.~Josefsson, ``Making architectural design phase obsolete - tdd as a design
	  method,''
	  \url{http://www.soberit.hut.fi/T-76.5650/Spring_2004/Papers/M.Josefsson_76650_final.pdf},
	  2004, t-76.650 Seminar course on SQA in Agile Software Development Helsinki
	  University of Technology. Último acesso em 01/03/2011.

@@

	\bibitem{mocks}
	C.~P. Mackinnon~T., Freeman~S., ``Endotesting: unit testing with mock
	  objects,'' in \emph{Extreme Programming Examined}, G.~Succi and M.~Marchesi,
	  Eds.\hskip 1em plus 0.5em minus 0.4em\relax Addison-Wesley Longman Publishing
	  Co., 2001, pp. 287--301.

	\bibitem{gof}
	e.~a. Eric T~Freeman, Elisabeth~Robson, \emph{Head First Design Patterns},
	  1st~ed.\hskip 1em plus 0.5em minus 0.4em\relax O'Reilly Media, 2004.

	\bibitem{feathers-synergy}
	M.~Feathers, ``The deep synergy between testability and good design,''
	  \url{http://michaelfeathers.typepad.com/michael_feathers_blog/2007/09/the-deep-synerg.html},
	  2007, Último acesso em 27/10/2010.

>>>>

% CHANGE #7
\begin{abstract}

	Despite the fact that Test-Driven Development (TDD) appears to be a software testing practice, 
	many developers affirm that the practice influences on class design. This study aims
	to better understand the effects of TDD and how the practice influences developers
	during class design on object-oriented systems. 
	We conduct an essentially qualitative exploratory study in which participants
	are invited to implement some exercises using TDD and, based on the data gathered,
	we raise details of how the practice influenced design decisions from the
	participants using interviews.
	We notice that the practice of TDD drives
	developers during class design by means of constant feedback about its quality. This study
	also names and catalogues feedback patterns perceived by the developers.

@@

% CHANGE #8 #9
When working on an experiment, researchers should decide between control
or realism. A realistic situation is usually
complex, which makes the comprehension of the phenomenon much
harder. On the other hand, increasing the control of the experiment reduces the
degree of realism, which can take the real factors of influence out of the study
scope \cite{guidelines-case-study}.

Based on the fact that the software development process involves many
human factors and is sensitive to the context that it belongs to,
we conducted an essentially qualitative exploratory study in which professionals
from the industry were invited to implement some prepared exercises using TDD and,
based on the gathered data, we investigated the details of how the practice influenced
the participants' class design decisions by means of interviews.

@@

It is a known fact that class design tends to lose quality during the software evolution \cite{evolution-lehman}.
However, it is hard to discuss quality in terms of class design. In this study context,
we used the SOLID principles, detailed in Martin's work \cite{bob-martin} and in the following
subsection.

%% ------------------------------------------------------------------------- %%
\subsection{SOLID principles}

@@

In Table \ref{tab:solid}, we relate the design smells to more common words
in object-oriented vocabulary. We also mention the principle that would solve
the problem. Through the rest of this work, we will make use of the mentioned
smells and SOLID principles.

\begin{table}
	\centering
	\begin{tabular}{| p{3cm} | p{4cm} | p{3cm} | }
		\hline
		\textbf{OO Concept} & \textbf{Design Smell} & \textbf{SOLID principle}\\

		Cohesion & Immobility, Viscosity & SRP, ISP \\ \hline

		Coupling & Fragility & OCP, LSP, DIP \\ \hline

		Encapsulation & Rigidity & SRP \\
		\hline
	\end{tabular}
	\caption{SOLID principles and smells}
	\label{tab:solid}
\end{table}

@@

Many empirical studies have been conducted to evaluate the effects of TDD.
In most of them, the effects of the practice in the class design are not taken
into account, and only the effect on external quality is measured.
Contrary to what this study proposes, many of them chose
to have a bigger control in the experiment. They were also conduted in academic
environments, with students that belong to different computer courses.

@@

However, studies that analyze TDD from the class design point of view do not
come to conclusive results; many of them even affirm that the effects of TDD
are not so different from those teams that do not practice it. Even the Janzen's PhD
thesis was inconclusive regarding the influence of the practice in coupling
and cohesion \cite{janzen-phd}.

@@

% CHANGE #3. It was already here, I mentioned again in the next section
Participants from different software development companies in the Brazilian
industry were selected. The participants' profiles are discussed in sub-section
\ref{sec:planejamento-participantes}. All of them were invited to implement
a few problems using Java in a defined timeframe. Participants practiced
TDD in one problem and did not practice in the other. The problems,
as well as the instruction of using or not TDD, were
randomized in order to reduce bias.

% CHANGE #34
All participants worked on the exercises in their offices, where they were
able to use their own computer and development IDE they are used to in 
their regular work. We were there with them at their offices, to explain
the mechanics and to reply to any doubt that would appear eventually.

@@

They evaluated the code based on three different categories: \textit{Simplicity}, \textit{Testability},
and \textit{Class Design Quality}. In each category, specialists scored from 1 (bad) to
5 (good) or chose not to evaluate that exercise. As a few participants did not finish the
implementation, specialists were warned to evaluate the intention of the participant and not
only the produced code.
To reduce the bias, the specialists did not know which code was written with TDD.

@@

% CHANGE #34, #31
We had 25 participants from 6 different companies. The participants, in their majority,
were not experienced in TDD. 10 of them affirmed to use TDD for no more than a year.
13 of them practiced TDD from 1 to 3 years. Only 1 practiced TDD from 3 to 4
years, and no participant was more experienced than that. Regardless of their level of experience,
all of them affirmed that they tried to practice their best TDD in the exercise they were supposed to.

% CHANGE #31
The numbers were a little different when talking about their experience in software
development. 6 of them worked with software development for the last 4 or 5 years.
7 of them worked from 6 to 10 years. Only 5 of them developed software for less
than 2 years. 16 of the participants use Java. However, 9 of them do not use
Java in their daily work. All of them know JUnit and 16 use mock objects 
\footnote{Mock objects are objects created during a unit test. They mock the behavior 
of a concrete object. Usually they are used to isolate the unit test from other classes. 
More information about it can be found on \cite{mocks}.}
during their development activities.
Only 3 have never heard about mock objects. When talking about object-orientation,
in the open question, most of the participants affirmed to have a good experience
on it. Some of them even affirmed to master the subject. Only a few of them affirmed
to have basic knowledge about OO.

@@

The elevated number of participants that do not use Java in their daily work is a consequence
of a company that uses PHP. However, we checked that, although they do not use Java constantly,
they did not have troubles during the implementation of the exercises.

% CHANGE #3
As in any experiment, one doubt that comes to mind is whether this sample is a good
generalization of the full population. However, in this (qualitative) study, generalization
is not our goal, but to deeply understand how TDD may help developers
during class design process. In this case, we believe that the selected sample will
help us understand it.

@@

We proposed four problems that should be implemented by the participants
\footnote{The exercises can be found at \url{https://gist.github.com/mauricioaniche/5694865}. Last access on June, the 1st, 2013.}. 
The goal was to simulate recurrent class design problems in several software projects.
In Table \ref{tab:problemas-exercicios}, we present the relation of a bad implementation of
the exercises and class design principles that were not followed.

@@

% CHANGE #21
When asking them to do their best in both
exercises, we intended to reduce the common bias that happens when you ask a participant to perform some task
using and not using the object in study. As it is "much more interesting to always use the new tool or practice",
asking them to be careful even when not using it, reduces the chance of the participants not taking care of
the code when not doing TDD.

\begin{table}
	\centering
	\begin{tabular}{| p{2cm} | p{5cm} | p{2.5cm} | }
		\hline
		\textbf{Exercise} & \textbf{Design Smell} & \textbf{Principles to}\\
		& & \textbf{be Followed}\\

@@

		Exercise 1 & Rigidity, Needless Complexity & SRP, OCP \\ \hline
		Exercise 2 & Fragility, Viscosity, Immobility & SRP, DIP, OCP \\ \hline
		Exercise 3 & Rigidity, Fragility & SRP\\
		Exercise 4 & Fragility, Viscosity, Immobility & OCP, SRP, DIP \\ \hline

@@

% CHANGE #3
As said before, each participant will work on two out of these four exercises.
In one of them, s/he will make use of TDD. In the other, s/he will not. In order to
reduce bias, the assignment of the exercises for each participant was randomized.

% CHANGE #2
Although the exercises seem small, most of participants thought they contained 
real world design issues. When asked about it during the questionnaire, a participant
said that \textit{"we frequently face problems like that, in which we have trouble to
define the responsibilities of each class"}. On the other hand, a few of them said
that real world problems tend to be more complex.

In our interpretation, the
number of lines needed to solve the problems is small when compared to real world systems,
but the exercises deal with complicated design problems, such as high coupling, low cohesion, and no
encapsulation. In order to create a decent solution, developers need to make good
use of object-oriented concepts. So, to sum up, we believe that these exercises will force
developers to be creative in terms of the solution.

@@

The statistical test used was the Wilcoxon. It is a non parametric hypothesis test
used to compare two samples and to check whether there is a difference in the meaning
of both samples. Therefore, we used Wilcoxon to compare the meaning between
the code metrics in code produced with and without TDD. The significance level
used was the standard (0.05). It means that we can reject the hypothesis if we
find a value smaller than that. Our hypotheses were the following:

% CHANGE #33
\begin{itemize}
	\item \textit{H1}. Code produced with the practice of TDD does not present a lower cyclomatic
	complexity than code produced without the practice of TDD.

	\item \textit{H2}. Code produced with the practice of TDD does not present a lower efferent
	coupling than code produced without the practice of TDD.

	\item \textit{H3}. Code produced with the practice of TDD does not present a higher cohesion
	of methods than code produced without the practice of TDD.

	\item \textit{H4}. Code produced with the practice of TDD does not present a lower number
	of lines per method than code produced without the practice of TDD.

	\item \textit{H5}. Code produced with the practice of TDD does not present a lower number of
	methods per class than code produced without the practice of TDD.
\end{itemize}

@@

In Table \ref{metricas-industria}, we show the \textit{p-values} found for
the differences between codes produced with and without TDD. By the numbers,
we observe that there was no significant differences in cyclomatic complexity
and efferent coupling.
Lack of cohesion of methods was different in two exercises (1 and 4). A difference
also appeared on the quantity of lines in a method (exercise 4) and quantity of
methods (exercise 1). However, it is not strong enough to affirm that there was
a difference in both quality attributes.

@@

\begin{table*}
	\centering
	\begin{tabular}{ | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{1.5cm} |}
		\hline
		\textbf{Exercise} & \textbf{Cyclomatic Complexity (H1)} & \textbf{Efferent coupling (H2)} & \textbf{Lack of Cohesion of Methods (H3)} & \textbf{Number of lines per method (H4)} 
		& \textbf{Quantity of methods per class (H5)} \\
		\hline
		Exercise 1 &	0.8967	&	0.6741 &	\cellcolor[gray]{0.8}2.04E-07* &	0.4962 &	\cellcolor[gray]{0.8}2.99E-06* \\
		Exercise 2	& 0.7868	&	0.7640 &	0.06132 &	0.9925 &	0.7501 \\
		Exercise 3	& 0.5463	&	0.9872 &	0.5471 &	0.7216 &	0.3972\\
		Exercise 4	& 0.2198	&	0.1361 &	\cellcolor[gray]{0.8}0.04891* &	\cellcolor[gray]{0.8}0.0032* &	0.9358\\
		\hline
		All together &	0.8123	&	0.5604 &	0.3278 &	0.06814 &	0.5849\\
		\hline
	\end{tabular}
	\caption{\textit{P-values} found for the difference between code produced with and without TDD}
	\label{metricas-industria}
\end{table*}

@@

%% ------------------------------------------------------------------------- %%
\subsection{Specialists' opinion}

% CHANGE #16
As mentioned before, we invited two software specialists to evaluate the produced by
all the participants. They did not know whether a piece of code was developed using
TDD. Both specialists have more than 10 years of experience in software development
and in object-oriented design, and act as technical leaders in their companies.

@@

The values presented above corroborate with many of the related work.
Apparently TDD does not influence the code to the point of changing
the metric values for coupling, cohesion, and simplicity. However, this is
incoherent to the industry's common feeling of the positive effects of TDD on class design.

% CHANGE #30
As stated, in this study we conducted a qualitative analysis to understand
how this influence goes from the point of view of the developers.
We chosed to interview those participants that have presented some difference between
their answers in the questionnaire and their source code. For instance, if a participant
mentioned that TDD helped him during class design, but his code did not present any improvement,
we then selected him to understand why it happened. The selection protocol was based on the fact that,
in qualitative studies, it is interesting to analyse "corner cases", as they can provide useful
and unexpected information.

@@

% CHANGE #18
Different from what was expected, almost all of the participants affirmed that the
practice of TDD would not change the class design produced by them during the exercises
in which they did not use the practice.
The main justification was that experience and previous knowledge on object-orientation
guided them during the class design. All participants agreed that a developer with no
knowledge in any of the mentioned areas would create a good class design only by practicing
TDD.

@@

% CHANGE #18
More than half of the participants also commented that one difference that is
perceived when they practice TDD is the constant feedback. In traditional testing,
the time between the production code writing and test code writing is too long.
When practicing TDD, developers are forced to write the test first, and receive all
feedback a test can provide sooner.

@@

% CHANGE #25
Also, the existence of the word \textit{"if"} in the name of the test is another
pattern. Tests that contain a name like that usually indicate the existence
of an "if" in the production code. The existence of the "if" by itself is not the problem;
but when we find a sequence of conditionals that can be replaced by the use of polymorphism,
those conditionals can be considered harmful in an object-oriented project. 
We call it \textbf{Conditional in Test's name}.

@@

		\textbf{Pattern} & \textbf{Class Design Smell} & \textbf{OO Principle to Fix The Problem}\\

@@

%% ------------------------------------------------------------------------- %%
% CHANGE #20, #22, 
\subsection{Relationship between Patterns and Unit Test Smells}

There are already some work on test code smells. In his book, Meszaros \cite{meszaros}
enumerates a couple of them. Many of his patterns match with the patterns we have
described. In Table \ref{tab:eu-e-meszaros}, we relate his patterns to ours.

As an example, he mentions a pattern called \textit{Obscure Test}, which is
a test that is difficult to understand at a glance. This one can be related to our
\textit{Too Big Scenario} pattern, as a big scenario makes the test harder to be
understood.

The \textit{Hard-To-Test Code} that, according to him, is a code that is difficult
to test, may happen because of a highly coupled code. He mentions that a possible solution
to this problem is to make use of stubs or mock objects. However, when making much use of it,
developers end up with our pattern \textit{Too Many Mock Objects} or even \textit{Non Used Mock Objects}.

Our patterns \textit{Same Change in Different Tests} and \textit{Repeated Tests for Different
Entities} can be related to his \textit{Test Code Duplication}, which is the same test code
repeated many times. He mentions that it can happen because of copy-and-paste code reuse but,
as we discussed, it also can happen because of a bad abstraction in the production code.

As we can see, all these test smells can be related to our patterns. In his work, Meszaros demonstrates
many causes for these smells, but we also argue that they can happen not only because the developer
was not paying attention to the test code quality, but because of a problem in the production code.
We reinforce that there is a strong connection between the quality of the test code and the quality
of the production code that is being tested.

\begin{table}[h!]
	\centering
	\begin{tabular}{| p{6cm} | p{6cm}| }
		\hline

		\textbf{Meszaros' Pattern} & \textbf{Test Feedback Pattern}\\

		Obscure Test & Too Big Scenario \\ \hline
		Hard-To-Test Code & Too Many Mock Objects, Non Used Mock Objects \\ \hline
		Test Code Duplication & Same Change in Different Tests, Repeated Tests for Different Entities \\
		\hline

	\end{tabular}
	\caption{Relationship between Meszaros' patterns and the patterns presented in this work}
	\label{tab:eu-e-meszaros}
\end{table}

@@

The proposed exercises were small compared to a real project. However, all exercises
contain localized class design problems. Once this study evaluates the effects
of TDD on class design, we believe the exercises can simulate real design problems
satisfactorily.
Besides, at the end of the implementation, participants answered in the questionnaire about the
similarity between the exercises and real world problems. All of them affirmed that
the problems found in the exercises were very similar to the ones they deal with in real world.

@@

% CHANGE #36, #26
As the participants were working on a fixed time box, some of them were not able to finish 
their implementations. This can influence the
quantitative analysis, as a design that would be complex at the end, looks simple
to an automated code metric. However, as they main contribution of this work was
the qualitative study, we do not see this problem as an important threat.

@@

%% ------------------------------------------------------------------------- %%
\section{Conclusion}
\label{sec:conclusion}

@@

In subsections below, we answer each of the questions raised by this study.

@@

The practice of TDD \textbf{can} affect the creating process of class design. However,
unlike the comments made by the industry, \textbf{the TDD practice cannot guide the
developer to a good class design automatically}; the experience and knowledge of
the developer are essential to develop object-oriented software.

@@

\textbf{Therefore, this is the way how the practice guides the developer to
a better class design; giving constant feedback about potential issues in
class design. It is a developer's task to realize these issues and 
improve the design. }

@@

%% ------------------------------------------------------------------------- %%
\begin{thebibliography}{38}

@@

	\bibitem{TDDByExample}
	K. Beck, \emph{Test-Driven Development By Example}, 1st~ed.\hskip 1em plus 0.5em
	  minus 0.4em\relax Addison-Wesley Professional, 2002.

	\bibitem{wambler-survey-agile}
	Ambler, S., ``How agile are you? 2010 survey results,''
	  \url{http://www.ambysoft.com/surveys/howAgileAreYou2010.html}, 2010, Last access on November, the 28th, 2010.

	\bibitem{versionone-2012}
	V.~One, ``State of agile development survey results,''
	  \url{http://www.versionone.com/state_of_agile_development_survey/11/}, 2012,
	  Last access on February, the 29th, 2012.

@@

	\bibitem{GOOS}
	S. Freeman., N. Pryce, \emph{Growing Object-Oriented Software, Guided by Tests},
	  1st~ed.\hskip 1em plus 0.5em minus 0.4em\relax Addison-Wesley Professional,
	  2009.

@@

	\bibitem{aniche-wbma}
	M.~Aniche, T.~Ferreira, and M.~Gerosa, ``What concerns beginner test-driven
	  development practitioners: A qualitative analysis of opinions in an agile
	  conference,'' 1st Brazilian Workshop on Agile Methods (WBMA), 2011.

@@

	\bibitem{langr}
	J.~Langr, ``Evolution of test and code via test-first design,''
	  \url{http://www.objectmentor.com}, 2001, Last access on March, the 1st, 2011..

@@

	\bibitem{meszaros}
	G.Mezsaros, ``xUnit Test Patterns: Refactoring Test Code,'' Addison-Wesley, 2007.

	\bibitem{josefsson}
	M.~Josefsson, ``Making architectural design phase obsolete - tdd as a design
	  method,''
	  \url{http://www.soberit.hut.fi/T-76.5650/Spring_2004/Papers/M.Josefsson_76650_final.pdf},
	  2004, t-76.650 Seminar course on SQA in Agile Software Development Helsinki
	  University of Technology. Last access on March, the 1st, 2011.

@@

	\bibitem{mocks}
	T. Mackinnon, S. Freeman, ``Endotesting: unit testing with mock
	  objects,'' in \emph{Extreme Programming Examined}, G.~Succi and M.~Marchesi,
	  Eds.\hskip 1em plus 0.5em minus 0.4em\relax Addison-Wesley Longman Publishing
	  Co., 2001, pp. 287--301.

	\bibitem{gof}
	E. Freeman, E. Freeman, B. Bates, K. Sierra, E. Robson, \emph{Head First Design Patterns},
	  1st~ed.\hskip 1em plus 0.5em minus 0.4em\relax O'Reilly Media, 2004.

	\bibitem{feathers-synergy}
	M.~Feathers, ``The deep synergy between testability and good design,''
	  \url{http://michaelfeathers.typepad.com/michael_feathers_blog/2007/09/the-deep-synerg.html},
	  2007, Last access on November, the 27th, 2012.
